---
title: "Lab5.Rmd"
author: "Jessica Atuahene"
date: "9/30/2020"
output: html_document
Bibliography: references.bib 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

```{r, confirmed_March}
Confirmed_State_6_13 <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/03-13-2020.csv")) %>%
  rename(Country_Region = "Country/Region", Province_State = "Province/State") %>% 
  filter (Country_Region == "US") %>% 
  group_by(Province_State, Country_Region) %>% 
  summarise(Confirmed = sum(Confirmed)) 

str(Confirmed_State_6_13)
```

```{r, confirmed_September}
Confirmed_State_9_13 <-   read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/09-13-2020.csv")) %>% 
  filter (Country_Region == "US") %>% 
  group_by(Province_State, Country_Region) %>% 
  summarise(Confirmed = sum(Confirmed)) 

str(Confirmed_State_9_13)
```

```{r}
## Notice from the above `str’ calls that there are 53 states in the table on 6/13/2020 and 58 states in the table on 9/13/2020

setdiff(Confirmed_State_9_13$Province_State, Confirmed_State_6_13$Province_State)
```

```{r,}
## There are 4 US territories a field you may have notice last week “Recovered”. We don’t need the “Recovered” row so let’s filter it from the data set
Confirmed_State_9_13 <- Confirmed_State_9_13 %>% 
  filter(Province_State != "Recovered") 
```

###Exercise 2 
```{r, exercise2,fig.width = 5, fig.height = 10 }
## Revise the above example for joining tables using 6/13/2020 and 9/13/2020. Plot the data as a bar plot. Make sure your graph looks nice in the report by adjusting the height of the graph in the R code chunk header.

Confirmed_State_6_13_9_13_joined <- full_join(Confirmed_State_6_13,
      Confirmed_State_9_13, by = c("Province_State"))
head(Confirmed_State_6_13_9_13_joined)

Confirmed_State_6_13_9_13_joined <- full_join(Confirmed_State_6_13,
      Confirmed_State_9_13, by = c("Province_State")) %>% 
      rename(Confirmed_6_13_2020 = "Confirmed.x", Confirmed_9_13_2020 = "Confirmed.y") %>% 
      select(-Country_Region.x, -Country_Region.y) %>% 
      replace_na(list(Confirmed_6_13_2020 = 0))
head(Confirmed_State_6_13_9_13_joined)

## which(is.na(Confirmed_State_6_13_9_13_joined))

Confirmed_State_6_13_9_13_joined_long <- Confirmed_State_6_13_9_13_joined %>% 
              pivot_longer(-c(Province_State),
                            names_to = "Date", values_to = "Confirmed")

ggplot(Confirmed_State_6_13_9_13_joined_long, aes(x = Confirmed,  y = Province_State))  + geom_bar(stat="identity",aes(color = Date))
```
### Exercise 3
```{r, Exercise3, fig.width = 5, fig.height = 10}
## Add more informative x and y labels and a title to the above graph in Ex2.
ggplot(Confirmed_State_6_13_9_13_joined_long, aes(x = Confirmed,  y = Province_State))  + geom_bar(stat="identity",aes(color = Date)) +  labs(title="COVID-19 Confirmed Cases in US",
        x ="Number of Confirmed Cases", y = "State/Province in US")
```

```{r}
library(lubridate)
```
 
### Exercise 3 
```{r,exercise5,fig.width = 5, fig.height = 10}
## Use Mutate to make a new column with deaths/confirmed cases (Hint: this might be best done after joining the tables, but before pivoting longer).

time_series_confirmed <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv")) %>%
  rename(Province_State = "Province/State", Country_Region = "Country/Region")

##  Have the date on your computer and then load the file into R.

download.file(url="https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv", 
               destfile = "data/time_series_covid19_confirmed_global.csv")

time_series_confirmed <- read_csv("data/time_series_covid19_confirmed_global.csv")%>%
  rename(Province_State = "Province/State", Country_Region = "Country/Region")

## Check the table properties to make sure the data imported as we expected
head(time_series_confirmed)

## As noted above this data is in wide format. To convert to long format
time_series_confirmed_long <- time_series_confirmed %>% 
               pivot_longer(-c(Province_State, Country_Region, Lat, Long),
                            names_to = "Date", values_to = "Confirmed") 

## Let’s look at the format of the data frame (tibble) now
head(time_series_confirmed_long)

## It would be convenient to have the confirmed cases and deaths in the same table. We can create another table with the deaths and then join the two tables
 download.file(url="https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv", 
               destfile = "data/time_series_covid19_deaths_global.csv")
 
 time_series_deaths <- read_csv("data/time_series_covid19_deaths_global.csv")%>%
  rename(Province_State = "Province/State", Country_Region = "Country/Region")

 ## As noted above this data is in wide format. To convert to long format
 
 time_series_deaths_long <- time_series_deaths %>% 
               pivot_longer(-c(Province_State, Country_Region, Lat, Long),
                            names_to = "Date", values_to = "Deaths") 
head(time_series_deaths_long)

## None of the columns in this data set meet that criteria, but we can create a column to serve as our key by mergering the names for the Province_State, Country_Region and Date columns using unite

time_series_confirmed_long <- time_series_confirmed_long %>% 
  unite(Key, Province_State, Country_Region, Date, sep = ".", remove = FALSE)
head(time_series_confirmed_long)

## For the second table we can do the same and eliminate the columns that are redundant so that we just have the Key and Deaths columns

time_series_deaths_long <- time_series_deaths_long %>% 
  unite(Key, Province_State, Country_Region, Date, sep = ".") %>% 
  select(Key, Deaths)

## Where there are not matching values, returns NA for the one missing. select can be used to remove the key after the join since we don’t have further need for it.
time_series_long_joined <- full_join(time_series_confirmed_long,
              time_series_deaths_long, by = c("Key")) %>% 
              select(-Key)
head(time_series_long_joined)

## Check to make sure neither the Cofirmed or Death counts have NA as a value
which(is.na(time_series_long_joined$Confirmed))


```{r}
##This results in the values for Guam, Northern Mariana Islands, Puerto Rico and the Virgin Islands being NA for 06/13

tail(Confirmed_State_6_13_9_13_joined, 5)
```


```{r}
## Let’s turn theNAs into 0. Also notice that we know have columns Confirmed.x Country_Region.x, Confirmed.y Country_Region.y. So let’s rename the Confirmed columns and delete the Country_Region columns

Confirmed_State_6_13_9_13_joined <- full_join(Confirmed_State_6_13,
      Confirmed_State_9_13, by = c("Province_State")) %>% 
      rename(Confirmed_6_13_2020 = "Confirmed.x", Confirmed_9_13_2020 = "Confirmed.y") %>% 
      select(-Country_Region.x, -Country_Region.y) %>% 
      replace_na(list(Confirmed_6_13_2020 = 0))
head(Confirmed_State_6_13_9_13_joined)

which(is.na(Confirmed_State_6_13_9_13_joined))
```

```{r}
##gather and spead were used in the Data Carpentries tutorial to switch between wide and long table formats. They work fine, but are being replaced by the newer functions pivot_longer and pivot_wider. which we will use below. See the pivot vignette for other examples.

Confirmed_State_6_13_9_13_joined_long <- Confirmed_State_6_13_9_13_joined %>% 
              pivot_longer(-c(Province_State),
                            names_to = "Date", values_to = "Confirmed")
```

```{r,fig.width = 5, fig.height = 10}
## Now we can use this table in plot that shows the confirmed cases for both dates.
# to adjust figure size {r, fig.width = 5, fig.height = 10}
ggplot(Confirmed_State_6_13_9_13_joined_long, aes(x = Confirmed,  y = Province_State))  + 
    geom_point(aes(color = Date)) 
```

```{r}
library(lubridate)
```

```{r} 
time_series_confirmed <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv")) %>%
  rename(Province_State = "Province/State", Country_Region = "Country/Region")

head(time_series_confirmed)

time_series_confirmed_long <- time_series_confirmed %>% 
               pivot_longer(-c(Province_State, Country_Region, Lat, Long),
                            names_to = "Date", values_to = "Confirmed") 

head(time_series_confirmed_long)

time_series_deaths <- read_csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv")%>%
  rename(Province_State = "Province/State", Country_Region = "Country/Region")

time_series_deaths_long <- time_series_deaths %>% 
               pivot_longer(-c(Province_State, Country_Region, Lat, Long),
                            names_to = "Date", values_to = "Deaths") 
head(time_series_deaths_long)

time_series_deaths_long <- time_series_deaths_long %>% 
  unite(Key, Province_State, Country_Region, Date, sep = ".") %>% 
  select(Key, Deaths)

time_series_confirmed_long <- time_series_confirmed_long %>% 
  unite(Key, Province_State, Country_Region, Date, sep = ".", remove = FALSE)
head(time_series_confirmed_long)

time_series_long_joined <- full_join(time_series_confirmed_long,
              time_series_deaths_long, by = c("Key")) %>% 
              select(-Key)
head(time_series_long_joined)

#which(is.na(time_series_long_joined$Confirmed))

#which(is.na(time_series_long_joined$Deaths))

time_series_long_joined$Date <- mdy(time_series_long_joined$Date)
``` 

###Exercise 4 
```{r, fig.width = 5, fig,height = 10}
##Using the time series data. Make a plot of the total number of confirmed deaths per day worldwide
time_series_long_joined %>% 
  group_by(Date) %>% 
  summarise_at(c("Confirmed", "Deaths"), sum) %>%
    ggplot(aes(x = Date,  y = Deaths)) + 
    geom_point() +
    geom_line() +
    ggtitle("Worldwide COVID-19 Deaths")
```

###Exercise 5
```{r, exercise 5, fig.width =5, fig.height = 10}
##Use Mutate to make a new column with deaths/confirmed cases (Hint: this might be best done after joining the tables, but before pivoting longer).

time_series_long_joined %>% 
  mutate(ratio = Deaths/Confirmed)
time_series_long_joined_counts <- time_series_long_joined %>% 
  pivot_longer(-c(Province_State, Country_Region, Lat, Long, Date, Deaths/Confirmed),
               names_to = "Report_Type", values_to = "Counts")
str(time_series_long_joined_counts)
```

###Exercise 6 
```{r,exercise6, fig.width = 5, fig.height = 10}
##Plot US deaths/confirmed cases per day.

time_series_long_joined %>% 
  group_by(Country_Region,Date) %>% 
  summarise_at(c("Confirmed", "Deaths"), sum) %>% 
  filter (Country_Region == "US") %>% 
    ggplot(aes(x = Date,  y = Deaths/Confirmed)) + 
    geom_point() +
    geom_line() +
    ggtitle("US COVID-19 Deaths")
```

###Exercise 7
```{r,exercise7, fig.width = 5, fig.height = 10}
##Make a single graph with from the times series data with the 10 countries with the hightest death total.

Confirmed <- time_series_long_joined %>% arrange(desc(Confirmed)) %>% slice(1:10)

time_series_long_joined %>% 
    group_by(Country_Region,Date) %>% 
    summarise_at(c("Confirmed", "Deaths"), sum) %>% 
    filter (Country_Region %in% c("Brazil","UK","Italy", 
                                "Mexico", "US","Spain","France","Russia","Germany","China")) %>% 
    ggplot(aes(x = Date,  y = Deaths, color = Country_Region)) + 
    geom_point() +
    geom_line() +
    ggtitle("COVID-19 Deaths")
```

###Exercise 8 
```{r, exercise8, fig.width = 5, fig.height = 10}
##Using Facet_Wraps make a graph from the times series data with the 10 countries with the hightest death total. Format the table to look nice in the report by changing the number of rows in your graph and/or adjusting the height in the R code chunk header.

time_series_long_joined %>% 
    group_by(Country_Region,Date) %>% 
    summarise_at(c("Confirmed", "Deaths"), sum) %>% 
    filter (Country_Region %in% c("Brazil","UK","Italy", 
                                "Mexico", "US","Spain","France","Russia","Germany","China")) %>% 
    ggplot(aes(x = Date,  y = Deaths, color = Country_Region)) + 
    geom_point() +
    geom_line() +
  facet_wrap(~Country_Region, ncol=2, scales="free_y") +
    ggtitle("COVID-19 Deaths")
``` 

###Exercise 9 
```{r, exercise9, fig.width = 5, fig.height = 10}
## Using Facet_Wraps make a graph from the times series data for confirmed cases for all US states and territories. Format the table to look nice in the report by changing the number of rows in your graph and/or adjusting the height in the R code chunk header.

download.file(url="https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_US.csv", 
               destfile = "data/time_series_covid19_confirmed_US.csv")

download.file(url="https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_US.csv", 
               destfile = "data/time_series_covid19_deaths_US.csv")
```

```{r}
time_series_confirmed_US <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_US.csv"))

head(time_series_confirmed_US)
```

```{r}
time_series_deaths_US <- read_csv(url("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_US.csv"))

head(time_series_deaths_US)
```
```{r}
time_series_confirmed_US_long <- time_series_confirmed_US %>% 
               pivot_longer(-c(UID, iso2, iso3, code3, FIPS, Admin2, Combined_Key, Province_State, Country_Region, Lat),
                            names_to = "Date", values_to = "Confirmed") 

head(time_series_confirmed_long)
```

```{r}
time_series_deaths_US_long <- time_series_deaths_US %>% 
               pivot_longer(-c(UID, iso2, iso3, code3, FIPS, Admin2, Combined_Key, Province_State, Country_Region, Lat),
                            names_to = "Date", values_to = "Deaths") 
head(time_series_deaths_long)
```

### Exercise 10 

```{r, exercise10, fig.width = 5, fig.height = 10}
##Apply a theme other than the one used in the Data Carpentries tutorial to the graph in Ex8.

theme_set(theme_gray())
time_series_long_joined %>% 
    group_by(Country_Region,Date) %>% 
    summarise_at(c("Confirmed", "Deaths"), sum) %>% 
    filter (Country_Region %in% c("Brazil","UK","Italy", 
                                "Mexico", "US","Spain","France","Russia","Germany","China")) %>% 
    ggplot(aes(x = Date,  y = Deaths, color = Country_Region)) +
    geom_point() +
    geom_line() + 
    ggtitle("COVID-19 Deaths") +
  facet_wrap(~Country_Region, ncol=2, scales="free_y") + theme_void() + theme(axis.text.x = element_text(colour = "grey20", size = 12, angle = 90, hjust = 0.5, vjust = 0.5),
                                                                              axis.text.y = element_text(colour = "grey20", size = 12),
                                                                              strip.text = element_text(face = "italic"),
                                                                              text = element_text(size = 16))
    
```
```

# REFERENCES 
```
@Misc{Chang2015,
  Title                    = {shiny: Web Application Framework for R. R package version 0.12.1},
  Author                   = {Chang, W. and Cheng, J. and Allaire, JJ. and Xie, Y. and McPherson, J. },
  Year                     = {2015},
  Type                     = {Computer Program},
  Url                      = {http://CRAN.R-project.org/package=shiny}
}
@Article{RCoreTeam,
  Title                    = {R: A Language and Environment for Statistical Computing},
  Author                   = {{R Core Team}},
  Year                     = {2015},
  Type                     = {Journal Article},
  Url                      = {http://www.R-project.org}
}
```
      
    

















